---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(ggplot2)
library(ggvd)
library(magrittr)
library(patchwork)
```

This package aims to simplify making 2-4 way Venn diagrams with ggplot2, 
whilst allowing for easy customisation. It builds upon the 
[ggvenn](https://github.com/yanlinlin82/ggvenn) and 
[ggVennDiagram](https://github.com/gaospecial/ggVennDiagram) packages but is
(in my opinion) simpler use.

In future I hope to adapt the 2-3 way Euler 
diagrams from 
[VennDiagram](https://cran.r-project.org/web/packages/VennDiagram/index.html) 
as well. I will not expand this package past 4 sets as at that point, it would
be best to use UpSet plots (see [UpSetR](https://github.com/hms-dbmi/UpSetR) or
[ComplexUpset](https://github.com/krassowski/complex-upset)).

## Dummy data

Generally I use Venn diagrams to compare lists of proteins, so here we create
a vector of fake proteins to use in this vignette.

```{r}
# Make some fake protein names
generate_protein_names <- function(n) {
  protein_letters <- do.call(
    paste0, 
    replicate(3, sample(LETTERS, n, replace = TRUE), simplify = FALSE)
  )
  paste0(protein_letters, sample.int(9, n, replace = TRUE))
}

# Must ensure protein name are unique!
proteins <- generate_protein_names(n = 2000)
```

Then we sample these proteins to create some sets to compare.

```{r}
# The input for prepare_venn should be a list
set.seed(42)
results <- list(
  "Sample 1" = sample(proteins, 1651),
  "Sample 2" = sample(proteins, 687),
  "Sample 3" = sample(proteins, 872),
  "Sample 4" = sample(proteins, 712)
)
```

## `prepare_venn()`

Starting with a list of numeric or characters vectors (i.e. a list of sets) to
compare, you must construct a data.frame of the appropriate format using
`prepare_venn()`. It is important that within each vector in the list, there
are no duplicated elements.

```{r}
sapply(results, function(x) any(duplicated(x)))
```

```{r}
prepare_venn(results)
```

## `geom_venn()`

## `count_venn()`
